---
title: "brr: an R package for Bayesian inference on the ratio of two Poisson rates"
author: "Stéphane Laurent"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


The `brr` package performs Bayesian inference on the rate ratio $\phi=\frac{\lambda}{\mu}$ of the *two Poisson samples model*  given by two independent observations 
$$\begin{cases}
x \sim {\cal P}(\lambda S) \\ 
y \sim{\cal P}(\mu T)
\end{cases}$$ 
where $\lambda$ and $\mu$ are 
the unknown incidence rates and $S$ and $T$ are the known observation-opportunity
sizes, or, for short, the sample sizes. $S$ and $T$ are also called the times at risk when they represent some durations.

This model can also be used in the case of two independent series of observations 
$(x_i)_{i=1}^m$ and $(y_i)_{i=1}^n$ assuming 
$$\begin{cases}
x_i \sim_{\text{i.i.d.}} {\cal P}(\lambda S_i) \\ 
y_i \sim_{\text{i.i.d.}} {\cal P}(\mu T_i)
\end{cases}$$ 
by setting $S=\sum S_i$ and $T= \sum T_i$, and by using the sufficient statistics 
$x=\sum x_i$ and $y=\sum y_i$ as the observations.

The `brr` package implements the *semi-conjugate family of prior distributions*. 
 Precisely, for positive numbers $a,b,c,d$ to be set by the user, the following independent prior distributions are assigned on $\mu$ and $\phi:=\lambda/\mu$:
$$
\mu \sim {\cal G}(a,b) \quad \text{and } \quad \phi \sim  \frac{T+b}{S} \times {\cal B}'(c,d),
$$
Then the joint posterior on $(\mu,\phi)$ is given by  
$$
(\mu \mid \phi,x,y) \sim {\cal G}(a+x+y, b + \phi S+T) 
\quad \text{and } \quad (\phi\mid x,y) \sim \frac{T+b}{S} \times {\cal B}'(c+x,a+d+y). 
$$

In particular :

- when $a=c=0.5$ and $b=d=0$, the prior is the reference prior, also called the *non-informative prior*;

* when $a,b>0$, $c=0.5$ and $d=0$, the prior is the semi-reference prior, 
also called the *semi-informative prior*, when the arbitrary 
Gamma prior distribution ${\cal G}(a,b)$ is assigned on $\mu$. 

### Setting parameters with `brr`

Use the `Brr` function to set the prior parameters, the sample sizes, and the observed counts. One can proceed step by step, for example below we start by supplying the parameters $a$ and $b$ of the prior Gamma distribution on $\mu$:

```{r}
library(brr)
model <- Brr(a=2,b=3)
summary(model)
```

Since $c$ and $d$ were not supplied, `brr` automatically considers the non-informative prior on $\phi$. Equivalently the same `brr` object can be defined by 
`Brr(a=2, b=3, c=NULL, d=NULL)` or `Brr(a=2, b=3, c=0.5, d=0)`. 

The `brr` object can be used itself to  update it by adding parameters, for example: 

```{r}
model <- model(c=3, d=3, S=100, T=100)
summary(model)
```

Now that $a$, $b$, $c$, $d$, $S$ and $T$ have been supplied, the user can play with all functions related to the prior distributions.  
For example, `dprior(model, "lambda", .)` evaluates the density of the prior distribution of $\lambda$. This is convenient to generate a graphic with the `curve` function:

```{r, fig.width=3, fig.height=2.5}
par(mar=c(4, 3, 1, 1))
curve(dprior(model, "lambda", x), from=0, to=2, axes=FALSE, xlab=expression(lambda), ylab=NA)
axis(1);axis(2)
```

The `brr` package also provides a way to generate a plot with automatic aesthetics:

```{r, fig.width=3, fig.height=2.5}
par(mar=c(4, 3, 1, 1))
plot(model, dprior(lambda))
axis(2)
```


### Posterior inference with `brr` 

Posterior inference is available after setting the observed counts $x$ and $y$:

```{r}
model <- model(x=14, y=20)
summary(model)
```

Estimates are provided by the `coef` function:

```{r}
coef(model)
```


Posterior credibility intervals are provided by the `confint` function:

```{r}
confint(model) 
```


Predictions are provided by the `predict` function after adding the sample sizes of the future experiment:

```{r}
model <- model(Snew=500, Tnew=500)
predict(model)
```

